---
title: "Additional Processing for Analysis"
output: html_document
date: "January 20, 2025"
---

# Purpose

This script exists because the data processing required for publishing our data is different from the processing required for ROMN's analyses.
This work is probably incomplete, but hopefully it's a good start on any additional processing of water chemistry data from the EDD/WQP format into the format used for analysis.

# Setup

First we load the required packages, functions, and data.
For now, this code loads the activities and results tables from a saved EDD, but ideally we would run this on data queried from WQP.

```{r packages, functions, and data}
# Packages
library(tidyverse) # data wrangling
library(readxl) # reading excel files
library(here) # relative file paths

# Functions
source("format_edd.R")

# Data
activities <- read_excel(here("data/output/ROMN_101823_edd.xlsx"), sheet = "Activities")
results <- read_excel(here("data/output/ROMN_101823_edd.xlsx"), sheet = "Results")
```

# Data Processing

Currently, the code repository includes one additional function to help prepare water quality data for analysis.
This function, called `censor_for_analysis()`, accomplishes the following:

- When a field blank has a detected concentration, `censor_for_analysis()` changes the corresponding environmental sample's Result_Detection_Condition to "Not Detected" if the environmental sample's concentration is less than twice the blank's and "Detected Not Quantified" if it is 2-5 times the blank's.
- It sets Result_Detection_Condition to "Detected Not Quantified" for observations with the "NFNSU" flag, which occurs when a dissolved fraction exceeds the concentration of its corresponding total fraction by more than the total fraction's MDL.
- It changes Result_Text to NA for any observations where the Result_Detection_Condition is not "Detected And Quantified" (accounts for all cases above).

```{r}
# Run function to change detection conditions & censor data
updated_results <- censor_for_analysis(results, activities)

# Inspect data 
updated_results
```
