---
title: "CCAL Processing Template"
output: html_document
date: "January 29, 2025"
---

# Purpose

This script serves as a guide for creating an EQuIS Electronic Data Deliverable (EDD) from an Excel spreadsheet provided by CCAL.
To use this template, make a copy of this file in the code directory of the water quality repository, replace the constants in the setup section, and press run!
Much of the process is automated, but after running the script you will need to manually review certain pieces as indicated later in this document.

# Setup

First, we define the constants that our code will refer to throughout this script.
Specifically, please supply the name of your preferred copy of the database, the relevant CCAL deliverable(s), and the limits table, and double check that they are in the appropriate file directories.

If processing a new year of data, you may need to revise the EndDate column of the limits table.
This is also a good time to confirm whether the detection limits have remained the same or changed and update the table as appropriate.
We maintain a ROMN version of the limits table, but the instructions in the imdccal package (https://github.com/nationalparkservice/imd-ccal) are useful for understanding the table's usage and purpose.

```{r constants}
db_filename <- "SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb" # must be in data/databases
ccal_deliverable_filename <- "ROMN_070224.xlsx" # must be in data/raw. May supply a vector of multiple file names if desired
limits_filename <- "limits.rds" # must be in data/raw
```

Now we load the packages, functions, and data required to run the processing code.
All packages except for `imdccal` are available on CRAN and can be downloaded with the `install.packages()` function.
The `imdccal` package is only available via GitHub, hence the additional steps for installation indicated in the code chunk below.

```{r packages, functions, and data}
# Load imdccal package
# Steps for installation are commented out
# install.packages("remotes")
# library(remotes) # to facilitate downloading packages from github
# remotes::install_github("nationalparkservice/imd-ccal") 
library(imdccal) # nps package for formatting CCAL data

# Load all other packages
library(tidyverse) # data wrangling
library(readxl) # reading excel files
library(lubridate) # working with dates
library(openxlsx) # exporting excel files
library(here) # relative file paths
library(fpCompare) # comparing floating point numbers
library(DBI) # working with databases
library(odbc) # working with databases

# Load functions from our script
source("format_edd.R")

# Load ROMN's limits table for CCAL data
limits <- readRDS(here('data', 'raw', limits_filename))
```

# Creating the EDD

## The automated stuff

Here we run the automated routine for creating the EDD, and we store it on our R environment.
Optionally, uncomment relevant code below in order to write the EDD to a file.
See "format_edd.R" for details on the functions called in this code chunk.

```{r}
# Create EDD and store in environment
edds <- format_edd_ccal(ccal_file_paths = here("data", "raw", ccal_deliverable_filename),
                        db_file_path = here("data", "databases", db_filename),
                        pre_2020 = FALSE,
                        limits = limits,
                        concat = FALSE) # if multiple file paths are supplied, optionally change concat to 
                                       # TRUE to create one EDD instead of one for each file name

# Write the EDD to file
# write_data(all_data = edds,
#            format = "xlsx",
#            destination_folder = here("data/output"),
#            overwrite = TRUE,
#            suffix = "_edd",
#            num_tables = 4)

# Alternatively create EDD and write to excel file in one step
# write_edd(ccal_file_paths = here("data", "raw", ccal_deliverable_filename),
#           db_file_path = here("data", "databases", db_filename),
#           pre_2020 = FALSE,
#           format = "xlsx", # "csv" is the other option
#           destination_folder = here("data/output"),
#           overwrite = TRUE,
#           limits = limits,
#           concat = FALSE) # if multiple file paths are supplied, optionally change concat to
#                           # TRUE to create one EDD instead of one for each file name
```

## The manual stuff

To facilitate manual review, we create a single table that contains enough information to understand the results of the QC process without inspecting the full EDD.
This should reduce the effort associated with navigating the multiple tables and many columns of the EDD.
Again, you can review this table in your R environment, or you can uncomment the relevant code below to write the snapshot(s) of the EDD(s) to an Excel file.

```{r}
# Create table that is easier to review
edd_snapshots <- lapply(edds, function(edd) {
  edd$Results %>%
    left_join(edd$Activities, by = "Activity_ID") %>%
    select(Activity_ID, Lab_Batch_ID, Activity_Type, Result_Type, Characteristic_Name, 
           Filtered_Fraction, Result_Text, Result_Detection_Condition, Result_Qualifier, 
           Method_Detection_Limit, Lower_Quantification_Limit, Source_Flags, Result_Comment)
})

names(edd_snapshots) <- names(edds)

# Optionally, write the EDD snapshot to a file
# lapply(1:length(edd_snapshots), function(i) {
#   write.xlsx(edd_snapshots[[i]],
#              here("data", "output",
#                   paste0(str_extract(names(edd_snapshots)[[i]], "^.*(?=\\.)"), "_edd_snapshot.xlsx")))
# })
```

There are some comments from the CCAL data that require manual review in order to determine if a flag is needed and where in the flag hierarchy the flag should be placed.
Below, we filter for any results that have comments that require review.
Please see the Source_Flags column to review those comments. 

```{r}
# The ecologist will manually apply flags based on Source_Flags
lapply(edd_snapshots, function(snapshot) {
  snapshot %>%
    filter(!is.na(Source_Flags))
})
```

There may also be textual comments from CCAL that have not been automatically matched with rows in the EDD.
We display any such comments below for the ecologist to review and possibly use to apply relevant QC flags.

```{r}
# Sometimes these comments have no lab numbers associated with them so I did not attempt to automate their inclusion in Source_Flags
# If output is NA, there are no other comments from CCAL to consider
getCCALData(here("data", "raw", ccal_deliverable_filename)) %>%
  lapply(function(deliverable) {
    deliverable$metadata$quest_results_comments
  })
```

Finally, the ecologist should apply any other relevant manual QC before submitting the EDD.
