---
title: "TestAmerica EDD"
output: html_document
date: "2025-02-12"
---

# Purpose

Drafting the EQuIS EDD creation process for TestAmerica data.
Currently using TestAmerica data and a prior deliverable from 2021 to help maintain the same processing outputs.

# Setup

```{r}
# Packages
library(tidyverse)
library(here)
library(readxl)
library(fpCompare)

# Functions
source("format_edd.R")
source("test_edd.R")
`%notin%` <- Negate(`%in%`)
```

```{r}
# Example EDD
edd_old_file_path <- here('data/EDD_examples/ROMN_WQ_EDD_TestAmerica_20220602.xlsx')
projects_old <- read_excel(edd_old_file_path, sheet = "Projects")
locations_old <- read_excel(edd_old_file_path, sheet = "Locations")
activities_old <- read_excel(edd_old_file_path, sheet = "Activities")
results_old <- read_excel(edd_old_file_path, sheet = "Results")

# TA deliverable file names
ta_filenames <- c("280-147837-1_TalStandard.csv",
                  "280-148304-1_TalStandard.csv",
                  "280-149036-1_TalStandard.csv",
                  "280-149967-1_TalStandard.csv",
                  "280-150396-1_TalStandard.csv",
                  "280-151067-1_TalStandard.csv",
                  "280-152639-1_TalStandard.csv",
                  "280-154540-1_TalStandard.csv",
                  "280-156396-1_TalStandard.csv")

# Read and merge TA deliverables, and remove unnecessary rows
ta_deliverable <- read_csv(here(paste0("data/raw/", ta_filenames))) %>%
  filter(`Sample Type Desc` %notin% c("Lab Control Sample", "Lab Control Sample Duplicate", 
                                      "Method Blank", "Matrix Spike", "Matrix Spike Duplicate"))

# Load database tables
dbo <- load_db(here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"))

# Connect to database
SEI_DB <- DBI::dbConnect(odbc::odbc(), .connection_string = paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};Dbq=", 
                                                                   here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb")))
# Load tables for testamerica
flag_description <- DBI::dbReadTable(SEI_DB,"IMPORT_water_lab_TestAmerica_Flag_Descriptions")
unit_crosswalk <- DBI::dbReadTable(SEI_DB, "Xwalk_Parameter_SampleFraction_Medium_Unit")
methods_crosswalk <- DBI::dbReadTable(SEI_DB, "IMPORT_Water_Lab_TestAmerica_Analytical_Methods")
# Disconnect from database
DBI::dbDisconnect(SEI_DB)
```

# Projects

STATUS: Perfectly matches the historical EDD.
TO-DO: This should be converted to a function and called by the relevant CCAL and TestAmerica functions as the process is identical.

```{r}
db_file_path <- here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb")

SEI_DB <- DBI::dbConnect(odbc::odbc(), .connection_string = paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};Dbq=", 
                                                                     db_file_path))
ta_projects <- DBI::dbReadTable(SEI_DB, "tbl_Projects_EDD") %>%
  rename(`#Org_Code` = X.Org_Code)

DBI::dbDisconnect(SEI_DB)
```

# Activities

STATUS: Matches the historical EDD with a few acceptable exceptions.
TO-DO: Attempt to merge as much of the activities formatting for CCAL and TestAmerica deliverables as possible into a function.

```{r}
ta_activities <- ta_deliverable %>%
  mutate(`#Org_Code` = "ROMN",
         Project_ID = paste(`#Org_Code`, "SEI", sep = "_"),
         Activity_Start_Date = format(mdy_hm(`Collection Date`), "%Y%m%d"),
         Custody_ID = `Client Sample ID`, 
         Custody_ID_Join = str_extract(`Client Sample ID`, "\\d{2}[A-Z]{3}\\d{7}"),
         Medium = if_else(Matrix == "Solid", "Sediment", Matrix)) %>%
  left_join(dbo$Joined, by = c("Custody_ID_Join" = "SampleNumber")) %>%
  mutate(Location_ID = paste(`#Org_Code`, str_extract(EventName, "(?i)[a-z]*_[a-z0-9]*"), sep = "_"),
         Activity_ID_wDups = paste(Location_ID, Activity_Start_Date, "T", Medium, sep = "_")) %>%
  select(`#Org_Code`, EventName, Project_ID, Location_ID, Activity_ID_wDups, Medium, Activity_Start_Date, Custody_ID, Notes) %>%
  unique() %>%
  rename(Additional_Location_Info = EventName,
         Activity_Comment = Notes) %>%
  mutate(Activity_Type = case_when(str_detect(Additional_Location_Info, "_FD_W$") ~ 
                                     "Quality Control Sample-Field Replicate",
                                   str_detect(Additional_Location_Info, "_FB_W$") ~
                                     "Quality Control Sample-Field Blank",
                                   TRUE ~ "Sample-Routine"),
         Medium_Subdivision = NA,
         Assemblage_Sampled_Name = NA,
         Activity_Start_Time = NA,
         Activity_Start_Time_Zone = NA,
         Activity_End_Date =  NA,
         Activity_End_Time = NA,
         Activity_End_Time_Zone = NA,
         Activity_Relative_Depth = NA,
         Activity_Depth = NA,
         Activity_Depth_Unit = NA,
         Activity_Upper_Depth = NA,
         Activity_Lower_Depth = NA, 
         Activity_Depth_Reference = NA,
         Activity_Sampler = NA,
         Activity_Recorder = NA,
         Activity_Conducting_Organization = NA,
         Station_Visit_Comment = NA,
         Collection_Method_ID = if_else(str_detect(Activity_Type, "[S|s]ample"), "ROMN_GRAB", NA),
         Collection_Equipment_Name = if_else(str_detect(Activity_Type, "[S|s]ample"), "Water bottle", NA),
         Collection_Equipment_Description = NA,
         Gear_Deployment = NA,
         Container_Type	= NA,
         Container_Color = NA,	
         Container_Size = NA,
         Container_Size_Unit = NA,
         Preparation_Method_ID = NA,
         Chemical_Preservative = NA,
         Thermal_Preservative = NA,
         Transport_Storage_Description = NA,	
         Activity_Group_ID = NA,
         Activity_Group_Name = NA,
         Activity_Group_Type = NA,
         Collection_Duration = NA,
         Collection_Duration_Unit = NA,
         Sampling_Component_Name = NA,
         Sampling_Component_Place_In_Series = NA,	
         Reach_Length = NA,
         Reach_Length_Unit = NA,	
         Reach_Width = NA,
         Reach_Width_Unit = NA,
         Pass_Count = NA,
         Net_Type = NA,
         Net_Surface_Area = NA,	
         Net_Surface_Area_Unit = NA,	
         Net_Mesh_Size = NA,
         Net_Mesh_Size_Unit = NA,
         Boat_Speed = NA,
         Boat_Speed_Unit = NA,	
         Current_Speed = NA,
         Current_Speed_Unit = NA,
         Toxicity_Test_Type = NA,
         Collection_Effort = NA,
         Collection_Effort_Gear_Type = NA,
         Org_Source_Act_ID = NA) %>%
  group_by(Activity_ID_wDups) %>%
  arrange(Custody_ID) %>%
  mutate(count = row_number() - 1,
         Activity_ID = paste(Activity_ID_wDups, count, sep = "_")) %>%
  ungroup() %>%
  select(`#Org_Code`, Project_ID, Location_ID, Activity_ID_wDups, Activity_ID, Activity_Type, Medium,
         Medium_Subdivision, Assemblage_Sampled_Name, Activity_Start_Date, Activity_Start_Time,
         Activity_Start_Time_Zone, Activity_End_Date, Activity_End_Time, Activity_End_Time_Zone,
         Activity_Relative_Depth, Activity_Depth, Activity_Depth_Unit, Activity_Upper_Depth,
         Activity_Lower_Depth, Activity_Depth_Reference, Additional_Location_Info, Activity_Sampler,
         Activity_Recorder, Custody_ID, Activity_Conducting_Organization, Station_Visit_Comment,
         Activity_Comment, Collection_Method_ID, Collection_Equipment_Name, 
         Collection_Equipment_Description, Gear_Deployment, Container_Type, Container_Color, 
         Container_Size, Container_Size_Unit, Preparation_Method_ID, Chemical_Preservative, 
         Thermal_Preservative, Transport_Storage_Description, Activity_Group_ID, Activity_Group_Name,
         Activity_Group_Type, Collection_Duration, Collection_Duration_Unit, Sampling_Component_Name,
         Sampling_Component_Place_In_Series, Reach_Length, Reach_Length_Unit, Reach_Width,
         Reach_Width_Unit, Pass_Count, Net_Type, Net_Surface_Area, Net_Surface_Area_Unit, Net_Mesh_Size,
         Net_Mesh_Size_Unit, Boat_Speed, Boat_Speed_Unit, Current_Speed, Current_Speed_Unit,
         Toxicity_Test_Type, Collection_Effort, Collection_Effort_Gear_Type, Org_Source_Act_ID)
```

# Locations

STATUS: Perfectly matches the historical EDD.
TO-DO: This should be converted to a function and called by the relevant CCAL and TestAmerica functions as the process is identical.

```{r}
ta_locations <- ta_activities %>%
  mutate(Park_Code = str_extract(Activity_ID, "(?<=_)[:alpha:]*(?=_)"),
         Location_ID = str_extract(Activity_ID, ".*(?=_\\d{8})"),
         Name = str_extract(Location_ID, "(?<=_).*")) %>%
  select(`#Org_Code`, Park_Code, Location_ID, Name) %>%
  unique() %>%
  left_join(dbo$Site, by = c("Name")) %>%
  rename("Location_Name" = "Alias") %>%
  mutate(Location_Type = "River/Stream", 
         Lat_Lon_Method = "GPS-Unspecified", 
         Lat_Lon_Datum = "NAD83") %>% 
  left_join(dbo$SitesMetadata, by = c("Name" = "SiteName")) %>%
  mutate(Source_Map_Scale_Numeric = NA,
         Lat_Lon_Accuracy = NA,
         Lat_Lon_Accuracy_Unit = NA,
         Location_Description = NA,
         Travel_Directions = NA,
         Location_Purpose = NA,
         Establishment_Date = NA) %>%
  rename("HUC8_Code" = "HUC8",
         "HUC12_Code" = "HUC12") %>%
  mutate(Alternate_Location_ID = NA,
         Alternate_Location_ID_Context = NA,
         Elevation = NA,
         Elevation_Unit = NA,
         Elevation_Method = NA,
         Elevation_Datum = NA,
         Elevation_Accuracy = NA,
         Elevation_Accuracy_Unit = NA,
         Country_Code = NA) %>%
  rename("State_Code" = "StateCode",
         "County_Code" = "CountyName") %>%
  mutate(Drainage_Area = NA,
         Drainage_Area_Unit = NA,
         Contributing_Area = NA,
         Contributing_Area_Unit = NA,
         Tribal_Land_Indicator = NA,
         Tribal_Land_Name = NA,
         Well_ID = NA,
         Well_Type = NA,
         Aquifer_Name = NA,
         Formation_Type = NA,
         Well_Hole_Depth = NA,
         Well_Hole_Depth_Unit = NA,
         Well_Status = NA,
         Public_Latitude = NA,
         Public_Longitude = NA) %>%
  select(`#Org_Code`, Park_Code, Location_ID, Location_Name, Location_Type, # order for EDD
         Latitude, Longitude, Lat_Lon_Method, Lat_Lon_Datum, Source_Map_Scale_Numeric,
         Lat_Lon_Accuracy, Lat_Lon_Accuracy_Unit, Location_Description, Travel_Directions,
         Location_Purpose,	Establishment_Date, HUC8_Code, HUC12_Code, 
         Alternate_Location_ID,	Alternate_Location_ID_Context,	Elevation,	Elevation_Unit,
         Elevation_Method,	Elevation_Datum,	Elevation_Accuracy,	Elevation_Accuracy_Unit,	Country_Code,
         State_Code,	County_Code,	Drainage_Area,	Drainage_Area_Unit,	Contributing_Area,
         Contributing_Area_Unit,	Tribal_Land_Indicator,	Tribal_Land_Name,	Well_ID,	Well_Type,
         Aquifer_Name,	Formation_Type,	Well_Hole_Depth,	Well_Hole_Depth_Unit,	Well_Status,
         Public_Latitude,	Public_Longitude)
```

# Results

STATUS: Not done
TO-DO: QC

```{r}
activities <- ta_activities

ta_results <- ta_deliverable %>%
  mutate(`#Org_Code` = "ROMN",
         Activity_Start_Date = format(mdy_hm(`Collection Date`), "%Y%m%d"),
         Custody_ID = `Client Sample ID`, 
         Custody_ID_Join = str_extract(`Client Sample ID`, "\\d{2}[A-Z]{3}\\d{7}"),
         Medium = if_else(Matrix == "Solid", "Sediment", Matrix)) %>%
  left_join(dbo$Joined, by = c("Custody_ID_Join" = "SampleNumber")) %>%
  mutate(Location_ID = paste(`#Org_Code`, str_extract(EventName, "(?i)[a-z]*_[a-z0-9]*"), sep = "_"),
         Activity_ID_wDups = paste(Location_ID, Activity_Start_Date, "T", Medium, sep = "_")) %>%
  inner_join(activities %>% select(Activity_ID_wDups, Custody_ID, Activity_ID),
             by = c("Activity_ID_wDups", "Custody_ID")) %>%
  mutate(Characteristic_Name = if_else(Analyte == "Percent Moisture", "Moisture content", Analyte),
         Method_Speciation = NA,
         Filtered_Fraction = if_else(`Prep Type` == "Total/NA", "Total", `Prep Type`)) %>%
  left_join(flag_description %>% select(-Flag),
            by = join_by(Flag == Flag_TestAmerica)) %>%
  left_join(unit_crosswalk,
             by = join_by(Unit == ResultValueUnits,
                          Medium == Medium,
                          Filtered_Fraction == SampleFraction,
                          Characteristic_Name == ParameterName)) %>%
  mutate(Result_Numeric = as.numeric(if_else(Result == "ND", NA, Result)),
         Result_Text = case_when(Conversion == "Yes" & Operation == "*" ~ Result_Numeric * Factor,
                                 Conversion == "Yes" & Operation == "/" ~ Result_Numeric / Factor,
                                 TRUE ~ Result_Numeric),
         Lab_Reported_Result = if_else(Result == "ND", "ND", as.character(Result_Text)),
         Method_Detection_Limit = case_when(Conversion == "Yes" & Operation == "*" ~ `Low Limit` * Factor,
                                            Conversion == "Yes" & Operation == "/" ~ `Low Limit` / Factor,
                                            TRUE ~ `Low Limit`),
         Lower_Quantification_Limit = case_when(Conversion == "Yes" & Operation == "*" ~ `High Limit` * Factor,
                                                Conversion == "Yes" & Operation == "/" ~ `High Limit` / Factor,
                                                TRUE ~ `High Limit`),
         Result_Text = if_else(Result_Text %<=% Method_Detection_Limit, NA, Result_Text),
         Result_Detection_Condition = if_else(is.na(Result_Text), "Not Detected", "Detected And Quantified"), 
         Result_Unit = FinalValueUnits,
         Result_Qualifier = case_when(Result_Text %<=% Lower_Quantification_Limit ~ "J-R",
                                      !is.na(Flag_Equis) ~ Flag_Equis,
                                      TRUE ~ NA),
         Result_Status = "Pre-Cert",
         Result_Type = if_else(Result_Qualifier == "J-R", "Estimated", "Actual", missing = "Actual"),
         Upper_Quantification_Limit = NA,
         Limit_Comment = "Detection Limit Type = MDL",
         Result_Comment = "",
         Temperature_Basis = NA,
         Statistical_Basis = NA,
         Time_Basis = NA,
         Weight_Basis = NA,
         Particle_Size_Basis = NA,
         Precision = NA,
         Bias = NA,
         Confidence_Interval = NA,
         Upper_Confidence_Limit = NA,
         Lower_Confidence_Limit = NA,
         Result_Sampling_Point_Name = NA,
         Result_Depth_Height_Measure = NA,
         Result_Depth_Height_Measure_Unit = NA,
         Result_Depth_Altitude_Reference_Point = NA,
         Characteristic_Join = str_to_lower(Characteristic_Name)) %>%
  left_join(methods_crosswalk %>%
              mutate(Analyte_Join = str_to_lower(Analyte)),
            by = join_by(Characteristic_Join == Analyte_Join,
                         Medium == Matrix,
                         Filtered_Fraction == PrepType,
                         `Analysis Method` == Analysis_Method)) %>%
  mutate(Analytical_Method_ID = EQuIS_Method,
         Analytical_Remark = NA,
         Lab_ID = "TESTAMC_AZ",
         Lab_Remark_Code = NA,
         Analysis_Start_Date = format(mdy_hm(`Analysis Date`), "%Y%m%d"),
         Analysis_Start_Time = format(mdy_hm(`Analysis Date`), "%H:%M:%S"),
         Analysis_Start_Time_Zone = NA,
         Lab_Accreditation_Indicator = NA,
         Lab_Accreditation_Authority_Name = NA) %>%
  rename(Lab_Batch_ID = `Lab Sample ID`) %>%
  mutate(Lab_Sample_Preparation_ID =  NA,
         Lab_Sample_Preparation_Start_Date = NA,
         Lab_Sample_Preparation_Start_Time = NA,
         Lab_Sample_Preparation_Start_Time_Zone = NA,
         Dilution_Factor = NA,
         Num_of_Replicates = NA,
         Data_Logger_Line_Name = NA,
         Biological_Intent = NA,
         Biological_Individual_ID = NA,
         Subject_Taxon = NA,
         Unidentified_Species_ID = NA,
         Tissue_Anatomy = NA,
         Group_Summary_Count_or_Weight = NA,
         Group_Summary_Count_or_Weight_Unit = NA,
         Cell_Form = NA,
         Cell_Shape = NA,
         Habit_Name_1 = NA,
         Habit_Name_2 = NA,
         Habit_Name_3 = NA,
         Voltinism = NA,
         Pollution_Tolerance = NA,
         Pollution_Tolerance_Scale = NA,
         Trophic_Level = NA,
         Functional_Feeding_Group_1 = NA,
         Functional_Feeding_Group_2 = NA,
         Functional_Feeding_Group_3 = NA,
         File_Name_or_Resource_ID = NA,
         Resource_Date = NA,
         Resource_Title_Name = NA,
         Resource_Creator_Name = NA,
         Resource_Publisher_Name = NA,
         Resource_Publication_Year = NA,
         Resource_Volume_Pages = NA,
         Resource_Subject_Text = NA,
         Frequency_Class_Descriptor_1 = NA,
         Frequency_Class_Bounds_Unit_1 = NA,
         Frequency_Class_Lower_Bound_1 = NA,
         Frequency_Class_Upper_Bound_1 = NA,
         Frequency_Class_Descriptor_2 = NA,
         Frequency_Class_Bounds_Unit_2 = NA,
         Frequency_Class_Lower_Bound_2 = NA,
         Frequency_Class_Upper_Bound_2 = NA,
         Frequency_Class_Descriptor_3 = NA,
         Frequency_Class_Bounds_Unit_3 = NA,
         Frequency_Class_Lower_Bound_3 = NA,
         Frequency_Class_Upper_Bound_3 = NA,
         Taxonomist_Accreditation_Indicator = NA,
         Taxonomist_Accreditation_Authority_Name = NA,
         Result_File_Name = NA,
         Reportable_Result = "Y",
         Source_Flags = Flag, #CHECK
         Logger_Standard_Difference = NA,
         Logger_Percent_Error = NA,
         Analytical_Method_ID_Context = NA,
         Predicted_Result = NA) %>%
  select(`#Org_Code`, Activity_ID, Characteristic_Name, Method_Speciation,  # ORDER TO MATCH EDD
         Filtered_Fraction, Result_Detection_Condition, Result_Text,
         Result_Unit, Result_Qualifier, Result_Status, Result_Type, Result_Comment,
         Method_Detection_Limit, Lower_Quantification_Limit, Upper_Quantification_Limit, Limit_Comment,
         Temperature_Basis, Statistical_Basis, Time_Basis, Weight_Basis, Particle_Size_Basis,
         Precision, Bias, Confidence_Interval, Upper_Confidence_Limit,
         Lower_Confidence_Limit, Result_Sampling_Point_Name, Result_Depth_Height_Measure,
         Result_Depth_Height_Measure_Unit, Result_Depth_Altitude_Reference_Point,
         Analytical_Method_ID, Analytical_Remark, Lab_ID, Lab_Remark_Code, Analysis_Start_Date,
         Analysis_Start_Time, Analysis_Start_Time_Zone, Lab_Accreditation_Indicator,
         Lab_Accreditation_Authority_Name, Lab_Batch_ID, Lab_Sample_Preparation_ID,
         Lab_Sample_Preparation_Start_Date, Lab_Sample_Preparation_Start_Time,
         Lab_Sample_Preparation_Start_Time_Zone, Dilution_Factor, Num_of_Replicates,
         Data_Logger_Line_Name, Biological_Intent, Biological_Individual_ID, Subject_Taxon,
         Unidentified_Species_ID, Tissue_Anatomy, Group_Summary_Count_or_Weight,
         Group_Summary_Count_or_Weight_Unit, Cell_Form, Cell_Shape, Habit_Name_1,
         Habit_Name_2, Habit_Name_3, Voltinism, Pollution_Tolerance, Pollution_Tolerance_Scale,
         Trophic_Level, Functional_Feeding_Group_1, Functional_Feeding_Group_2,
         Functional_Feeding_Group_3, File_Name_or_Resource_ID, Resource_Date,
         Resource_Title_Name, Resource_Creator_Name, Resource_Publisher_Name,
         Resource_Publication_Year, Resource_Volume_Pages, Resource_Subject_Text,
         Frequency_Class_Descriptor_1, Frequency_Class_Bounds_Unit_1,
         Frequency_Class_Lower_Bound_1, Frequency_Class_Upper_Bound_1,
         Frequency_Class_Descriptor_2, Frequency_Class_Bounds_Unit_2,
         Frequency_Class_Lower_Bound_2, Frequency_Class_Upper_Bound_2,
         Frequency_Class_Descriptor_3, Frequency_Class_Bounds_Unit_3,
         Frequency_Class_Lower_Bound_3, Frequency_Class_Upper_Bound_3,
         Taxonomist_Accreditation_Indicator, Taxonomist_Accreditation_Authority_Name,
         Result_File_Name, Lab_Reported_Result, Reportable_Result, Source_Flags,
         Logger_Standard_Difference, Logger_Percent_Error, Analytical_Method_ID_Context,
         Predicted_Result)
```

# Testing

Note that we manually adjust some of the Activity_ID's for the purposes of comparison.
These differences occurred due to a discrepancy in the date (possibly adjusted manually during QC).

```{r}
comparison <- compare_edd(edd_old_file_path <- here('data/EDD_examples/ROMN_WQ_EDD_TestAmerica_20220602.xlsx'),
                          edd_new = list(Projects = ta_projects,
                                         Locations = ta_locations,
                                         Activities = ta_activities %>%
                                           mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_0" ~ 
                                                                            "ROMN_GRKO_S01_20210802_T_Sediment_0",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_1" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Sediment_1",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_0" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_0",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_1" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_1",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_2" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_2",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_3" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_3",
                                                                          TRUE ~ Activity_ID)),
                                         Results = ta_results %>%
                                           mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_0" ~ 
                                                                            "ROMN_GRKO_S01_20210802_T_Sediment_0",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_1" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Sediment_1",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_0" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_0",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_1" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_1",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_2" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_2",
                                                                          Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_3" ~
                                                                            "ROMN_GRKO_S01_20210802_T_Water_3",
                                                                          TRUE ~ Activity_ID))))
```

Summary/explanation of the differences:

- Activities:
    - Activity_Type (6): we are now correctly recording when something is a field blank or field dup rather than saying all are sample-routine
    - Activity_Start_Date (6): there are 6 observations where the date was historically 20210802 but in the new EDD is 20210803. I manually corrected the date part of the Activity_ID column in order to get my comparison function to work, but I left the date column unchanged. I suspect that these dates were corrected to 20210802 in the historical EDD as part of a manual QC process.
    - Activity_Comment (3): the differences appear to be due to incomplete sentences at the end of the comments in the old EDD.
    
- Results:
    - Result_Detection_Condition (2): I verified these two differences in the TestAmerica deliverable. They reported these as "ND" for "Not Detected", as the new EDD correctly states. The old EDD reported these as "Detected and Quantified".
    - Result_Text (2): I verified these two differences in the TestAmerica deliverable. They reported these as "ND" for "Not Detected", and the new EDD correctly censors these results. The old EDD reported these as equal to the MDL (we are also supposed to censor measurements equal to the MDL anyway).
    - Result_Unit (5): These 5 differences are in the unit of Moisture content. Unsure how to resolve; need to ask water quality party.
    - Result_Qualifier (273): ??
    - Result_Type (113): in the new EDD, we set the `Result_Type` to "Estimated" when the "J-R" flag is raised.
    - Result_Comment (646): ??
    - Lower_Quantification_Limit (112): These 112 differences are records that are NA in the old EDD but have an LQL in the new EDD. These are all observations that have the J-R flag -- it seems the VBA code set the LQL to NA when J-R was raised (but not sure where that happens). We should always include the LQL when possible.
    - Analysis_Start_Date (646): new code uses analysis start date (Analysis Date in TestAmerica deliverable) rather than activity start date
    - Analysis_Start_Time (646): new code uses analysis start time (Analysis Date in TestAmerica deliverable) instead of setting everything to midnight
    - Lab_Reported_Result (268): new code sets this to be exactly the result from the lab rather than using the MDL when a result is not detected and/or imputations
    - Reportable_Result (271): new code always sets this to 'Y'
