---
title: "TestAmerica_processing_template"
output: html_document
date: "2025-04-02"
---

# Purpose

This script serves as a guide for creating an EQuIS Electronic Data Deliverable (EDD) from CSV files provided by Test America.
To use this template, make a copy of this file in the code directory of the water quality repository, replace the constants in the setup section, and press run!

# Setup

First, we define the constants that our code will refer to throughout this script.
Specifically, please supply the name of your preferred copy of the database and the relevant Test America deliverable(s) and double check that they are in the appropriate file directories.

```{r constants}
db_filename <- "SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb" # must be in data/databases
ta_filenames <- c("280-193566-1_TalStandard.csv",
                  "280-191667-1_TalStandard.csv") # must be in data/raw. Recommended to supply all CSVs from a given field season
```

Now we load the packages, functions, and data required to run the processing code.
All packages except for `imdccal` are available on CRAN and can be downloaded with the `install.packages()` function.
The `imdccal` package is only available via GitHub, hence the additional steps for installation indicated in the code chunk below.

```{r packages, functions, and data}
# Load imdccal package
# Steps for installation are commented out
# install.packages("remotes")
# library(remotes) # to facilitate downloading packages from github
# remotes::install_github("nationalparkservice/imd-ccal") 
library(imdccal) # nps package for formatting CCAL data

# Load all other packages
library(tidyverse) # data wrangling
library(readxl) # reading excel files
library(lubridate) # working with dates
library(openxlsx) # exporting excel files
library(here) # relative file paths
library(fpCompare) # comparing floating point numbers
library(DBI) # working with databases
library(odbc) # working with databases

# Load functions from our script
source("format_edd.R")
```

# Creating the EDD

## The automated stuff

Here we run the automated routine for creating the EDD, and we store it on our R environment.
Optionally, uncomment relevant code below in order to write the EDD to a file.
See "format_edd.R" for details on the functions called in this code chunk.

```{r}
# Create EDD and store in environment
edd <- format_edd_test_america(test_america_file_paths = here(paste0("data/raw/", ta_filenames)), 
                               db_file_path = here("data", "databases", db_filename))

# Write the EDD from R environment to file
# The name of the file will be the name of the item in the list (ROMN_TestAmerica_2024) and the suffix (_edd) pasted together
# write_ccal(all_data = list(ROMN_TestAmerica_2024 = edd),
#            format = "xlsx",
#            destination_folder = here("data/output"),
#            overwrite = TRUE,
#            suffix = "_edd",
#            num_tables = 4)

# Alternatively create EDD and write to excel file in one step
# The name of the file will be "ROMN_TestAmerica", the year of the field season in the deliverables, and "_edd" pasted together
# write_edd(deliverable_file_paths = here(paste0("data/raw/", ta_filenames)),
#           db_file_path = here("data", "databases", db_filename),
#           pre_2020 = FALSE, # If you are using data from before 2020, set this to a file path to ROMN_SEI_in_EQuIS.xlsx, which includes a crosswalk between EventName and SampleNumber.
#           lab = "Test America",
#           format = "xlsx",
#           destination_folder = here("data/output"),
#           overwrite = TRUE)
```

## The manual stuff

To facilitate manual review, we create a single table that contains enough information to understand the results of the QC process without inspecting the full EDD.
This should reduce the effort associated with navigating the multiple tables and many columns of the EDD.
Again, you can review this table in your R environment, or you can uncomment the relevant code below to write the snapshot(s) of the EDD(s) to an Excel file.

```{r}
# Create table that is easier to review
edd_snapshot <- edd$Results %>%
    left_join(edd$Activities, by = "Activity_ID") %>%
    select(Activity_ID, Lab_Batch_ID, Activity_Type, Result_Type, Characteristic_Name, 
           Filtered_Fraction, Result_Text, Result_Detection_Condition, Result_Qualifier, 
           Method_Detection_Limit, Lower_Quantification_Limit, Source_Flags, Result_Comment)

# Optionally, write the EDD snapshot to a file
# The file name can be modified below
# write.xlsx(edd_snapshot,
#            here("data", "output", "ROMN_TestAmerica_2024_edd_snapshot.xlsx"))
```

Finally, the ecologist should review the data, apply any other relevant manual QC, update the Result_Status, and the EDD should be checked using the EQuIS Data Processor (EDP) before it is submitted.
