---
title: "Comparing the Output of Historical and Current Methods for Creating the EQuIS EDD"
output: html_document
date: "2025-02-12"
---

# Purpose

The purpose of this script is to use our current methods for creating EDDs from Test America deliverables to recreate historical EDDs, and to compare these new versions to what we originally created using the VBA routine in Access.
Note that the EDDs we recreate here are not fully complete; our streamlined EDD creation routine still requires some degree of manual review.

# Setup

First, we load the packages, functions, and data required to run this script.

```{r}
# Packages
library(tidyverse) # data wrangling
library(readxl) # reading excel files
library(openxlsx) # exporting excel files
library(here) # relative file paths
library(fpCompare) # comparing floating point numbers
library(DBI) # working with databases
library(odbc) # working with databases
library(imdccal) # has a function for writing EDDs

# Functions
source("format_edd.R")
source("test_edd.R")
```

# Create 2021 EDD

In this section we create an EDD for the CCAL data from 2021.
By default we store the EDD in the R environment, but you may uncomment code below to write it to an Excel file instead.

```{r}
# Define TA deliverable file names
ta_filenames_2021 <- c("280-147837-1_TalStandard.csv",
                  "280-148304-1_TalStandard.csv",
                  "280-149036-1_TalStandard.csv",
                  "280-149967-1_TalStandard.csv",
                  "280-150396-1_TalStandard.csv",
                  "280-151067-1_TalStandard.csv",
                  "280-152639-1_TalStandard.csv",
                  "280-154540-1_TalStandard.csv",
                  "280-156396-1_TalStandard.csv")

# Define TA deliverable file paths
ta_filepaths_2021 <- here(paste0("data/raw/", ta_filenames_2021))

# Create EDD
edd_2021 <- format_edd_test_america(test_america_file_paths = ta_filepaths_2021, 
                                    db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"))

# Write the EDD from R environment to file
# write_ccal(all_data = list(ROMN_TestAmerica_2021 = edd_2021),
#            format = "xlsx",
#            destination_folder = here("data/output"),
#            overwrite = TRUE,
#            suffix = "_edd",
#            num_tables = 4)

# Alternatively create EDD and write to excel file in one step
# write_edd(deliverable_file_paths = ta_filepaths_2021,
#           db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"),
#           pre_2020 = FALSE,
#           lab = "Test America",
#           format = "xlsx",
#           destination_folder = here("data/output"),
#           overwrite = TRUE)
```

Here, we create a single table that contains enough information to understand the results of the QC process without inspecting the full 2019 EDD across multiple tables.
If we were actually creating and submitting this EDD, the ecologist would need to review the lab comments in the Result_Comment column to determine whether additional flags are needed.

```{r}
# Create data that is easier to review
edd_2021_snapshot <- edd_2021$Results %>%
    left_join(edd_2021$Activities, by = "Activity_ID") %>%
    select(Activity_ID, Lab_Batch_ID, Activity_Type, Result_Type, Characteristic_Name, 
           Filtered_Fraction, Result_Text, Result_Detection_Condition, Result_Qualifier, 
           Method_Detection_Limit, Lower_Quantification_Limit, Source_Flags, Result_Comment)
```

# Test 2021 EDD

In this section, we compare the output from the EDD we just created for our 2021 data to the EDD that was created with our older process.

Note that we manually adjust some of the Activity_ID's for the purposes of comparison.
These differences occurred due to a discrepancy in the date (possibly adjusted manually during QC).

```{r}
comparison_2021 <- compare_edd(edd_old_file_path <- here('data/EDD_examples/ROMN_WQ_EDD_TestAmerica_20220602.xlsx'),
                               edd_new = list(Projects = edd_2021$Projects,
                                              Locations = edd_2021$Locations,
                                              Activities = edd_2021$Activities %>%
                                                mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_0" ~ 
                                                                                 "ROMN_GRKO_S01_20210802_T_Sediment_0",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_1" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Sediment_1",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_0" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_0",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_1" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_1",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_2" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_2",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_3" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_3",
                                                                               TRUE ~ Activity_ID)),
                                              Results = edd_2021$Results %>%
                                                mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_0" ~ 
                                                                                 "ROMN_GRKO_S01_20210802_T_Sediment_0",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Sediment_1" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Sediment_1",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_0" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_0",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_1" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_1",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_2" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_2",
                                                                               Activity_ID == "ROMN_GRKO_S01_20210803_T_Water_3" ~
                                                                                 "ROMN_GRKO_S01_20210802_T_Water_3",
                                                                               TRUE ~ Activity_ID))))
```

Summary/explanation of the differences:

- Activities:
    - Activity_Type (6): we are now correctly recording when something is a field blank or field dup rather than saying all are sample-routine
    - Activity_Start_Date (6): there are 6 observations where the date was historically 20210802 but in the new EDD is 20210803. I manually corrected the date part of the Activity_ID column in order to get my comparison function to work, but I left the date column unchanged. I suspect that these dates were corrected to 20210802 in the historical EDD as part of a manual QC process.
    - Activity_Comment (3): the differences appear to be due to incomplete sentences at the end of the comments in the old EDD.
    
- Results:
    - Result_Detection_Condition (114): Most of these differences occur because we now set the detection condition to "Present Below Quantification Limit" when a value is less than the LQL, but greater than or equal to the MDL. I verified the final two differences in the TestAmerica deliverable. They reported these as "ND" for "Not Detected", as the new EDD correctly states. The old EDD reported these as "Detected and Quantified".
    - Result_Text (111): We now censor all results less than the LQL
    - Result_Qualifier (456): we overhauled our flagging process
    - Result_Comment (516): we changed the structure of this field
    - Lower_Quantification_Limit (112): These 112 differences are records that are NA in the old EDD but have an LQL in the new EDD. These are all observations that have the J-R flag -- it seems the VBA code set the LQL to NA when J-R was raised (but not sure where that happens). We should always include the LQL when possible.
    - Lab_ID (646): Dean created a new option for us, EUROFINS_DEN, which we will use going forwards
    - Analysis_Start_Date (646): new code uses analysis start date (Analysis Date in TestAmerica deliverable) rather than activity start date
    - Analysis_Start_Time (646): new code uses analysis start time (Analysis Date in TestAmerica deliverable) instead of setting everything to midnight
    - Lab_Reported_Result (268): new code sets this to be exactly the result from the lab rather than using the MDL when a result is not detected and/or imputations
    - Reportable_Result (271): new code always sets this to 'Y'

# Create 2022 EDD

In this section we create an EDD for the CCAL data from 2022.

```{r}
# Define TA deliverable file names
ta_filenames_2022 <- c("280-166181-1_TalStandard.csv",
                       "280-166195-1_TalStandard.csv",
                       "280-166196-1_TalStandard.csv",
                       "280-168778-1_TalStandard.csv",
                       "280-164074-1_TalStandard.csv")

# Define TA deliverable file paths
ta_filepaths_2022 <- here(paste0("data/raw/", ta_filenames_2022))

# Create EDD
edd_2022 <- format_edd_test_america(test_america_file_paths = ta_filepaths_2022, 
                                    db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"))

# Write the EDD from R environment to file
# write_ccal(all_data = list(ROMN_TestAmerica_2022 = edd_2022),
#            format = "xlsx",
#            destination_folder = here("data/output"),
#            overwrite = TRUE,
#            suffix = "_edd",
#            num_tables = 4)

# Alternatively create EDD and write to excel file in one step
# write_edd(deliverable_file_paths = ta_filepaths_2022,
#           db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"),
#           pre_2020 = FALSE,
#           lab = "Test America",
#           format = "xlsx",
#           destination_folder = here("data/output"),
#           overwrite = TRUE)
```

Also see the EDD snapshot for easier manual review.

```{r}
edd_2022_snapshot <- edd_2022$Results %>%
    left_join(edd_2022$Activities, by = "Activity_ID") %>%
    select(Activity_ID, Lab_Batch_ID, Activity_Type, Result_Type, Characteristic_Name, 
           Filtered_Fraction, Result_Text, Result_Detection_Condition, Result_Qualifier, 
           Method_Detection_Limit, Lower_Quantification_Limit, Source_Flags, Result_Comment)
```

# Test 2022 EDD

In this section, we compare the output from the EDD we just created for our 2022 data to the EDD that was created with our older process.

Note that we manually adjust some of the Activity_ID's for the purposes of comparison.
These differences occurred due to a discrepancy in the date (possibly adjusted manually during QC).

```{r}
comparison_2022 <- compare_edd(edd_old_file_path <- here('data/EDD_examples/ROMN_WQ_EDD_20230309_PostEDP.xlsx'),
                               edd_new = list(Projects = edd_2022$Projects,
                                              Locations = edd_2022$Locations,
                                              Activities = edd_2022$Activities %>%
                                                mutate(Activity_ID = if_else(Activity_ID == "ROMN_GLAC_S04_20220719_T_Sediment_0",
                                                                             "ROMN_GLAC_S04_20220714_T_Sediment_0",
                                                                             Activity_ID)),
                                              Results = edd_2022$Results %>%
                                                mutate(Activity_ID = if_else(Activity_ID == "ROMN_GLAC_S04_20220719_T_Sediment_0",
                                                                             "ROMN_GLAC_S04_20220714_T_Sediment_0",
                                                                             Activity_ID))))
```

Summary/explanation of the differences:

- Projects:
    - Project_Start_Date (1): different date format
    
- Activities:
    - Activity_Type (5): we are now correctly recording when something is a field blank or field dup rather than saying all are sample-routine
    - Activity_Start_Date (70): The new EDD does not include the slashes in dates. In one case, I also manually corrected the date part of the Activity_ID column in order to get my comparison function to work, but I left the date column unchanged. I suspect that these dates were corrected in the historical EDD as part of a manual QC process.

- Results:
    - Result_Detection_Condition (127): Most of these differences occur because we now set the detection condition to "Present Below Quantification Limit" when a value is less than the LQL, but greater than or equal to the MDL. There is one observation which the old EDD reported these as "Detected and Quantified" but is actually a non-detect from the lab
    - Result_Text (127): we now censor all results less than the LQL
    - Result_Qualifier (432): we overhauled our flagging process
    - Result_Comment (486): we changed the structure of this field
    - Lower_Quantification_Limit (129): These differences are records that are NA in the old EDD but have an LQL in the new EDD. These are all observations that have the J-R flag -- it seems the VBA code set the LQL to NA when J-R was raised (but not sure where that happens). We should always include the LQL when possible.
    - Lab_ID (608): Dean created a new option for us, EUROFINS_DEN, which we will use going forwards
    - Analysis_Start_Date (608): new code uses analysis start date (Analysis Date in TestAmerica deliverable) rather than activity start date
    - Analysis_Start_Time (608): new code uses analysis start time (Analysis Date in TestAmerica deliverable) instead of setting everything to midnight
    - Lab_Reported_Result (253): new code sets this to be exactly the result from the lab rather than using the MDL when a result is not detected and/or imputations
    - Reportable_Result (253): new code always sets this to 'Y'

# Create 2019 EDD

In this section we create an EDD for the CCAL data from 2019.

THIS IS NOT WORKING!
It seems that analogous dissolved vs total fractions were not always collected or have somehow been lost...

```{r}
# Define TA deliverable file names
ta_filenames_2019 <- c("280-123562-1_TalStandard.csv",
                       "280-123566-1_TalStandard.csv",
                       "280-124111-1_TalStandard.csv",
                       "280-124119-1_TalStandard.csv",
                       "280-124141-1_TalStandard.csv",
                       "280-124485-1_TalStandard.csv",
                       "280-125452-1_TalStandard.csv",
                       "280-126433-1_TalStandard.csv",
                       "280-126425-1_TalStandard.csv",
                       "280-125633-1_TalStandard.csv",
                       "280-127663-1_TalStandard.csv",
                       "280-127918-1_TalStandard.csv",
                       "280-128955-1_TalStandard.csv",
                       "280-129879-1_TalStandard.csv",
                       "280-129898-1_TalStandard.csv",
                       "280-131288-1_TalStandard.csv",
                       "280-132126-1_TalStandard.csv")

# Define TA deliverable file paths
ta_filepaths_2019 <- here(paste0("data/raw/", ta_filenames_2019))

# Create EDD
edd_2019 <- format_edd_test_america(test_america_file_paths = ta_filepaths_2019, 
                                    db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"),
                                    pre_2020 = here("data/EDD_examples/ROMN_SEI_in_EQuIS.xlsx"))

# Write the EDD from R environment to file
# write_ccal(all_data = list(ROMN_TestAmerica_2019 = edd_2019),
#            format = "xlsx",
#            destination_folder = here("data/output"),
#            overwrite = TRUE,
#            suffix = "_edd",
#            num_tables = 4)

# Alternatively create EDD and write to excel file in one step
# write_edd(deliverable_file_paths = ta_filepaths_2019,
#           db_file_path = here("data/databases/SEI_ROMN_WQLab_Processing_20250131_LSmithCopy.accdb"),
#           pre_2020 = here("data/EDD_examples/ROMN_SEI_in_EQuIS.xlsx"),
#           lab = "Test America",
#           format = "xlsx",
#           destination_folder = here("data/output"),
#           overwrite = TRUE)
```

# Test 2019 EDD

This also won't work anymore but I was drafting it before I implemented the tot_vs_dissolved() function...

```{r}
comparison_2019 <- compare_edd(edd_old_file_path <- here('data/EDD_examples/ROMN_WQ_EDD_20200205.xlsx'),
                               edd_new = list(Projects = edd_2019$Projects,
                                              Locations = edd_2019$Locations,
                                              Activities = edd_2019$Activities %>%
                                                mutate(Activity_ID = str_replace(Activity_ID, "_T_", "_")) %>%
                                                mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20190326_Water_0" ~ 
                                                                                 "ROMN_GRKO_S01_20190526_Water_0", # date adjustment
                                                                               Activity_ID == "ROMN_GRKO_S01_20190326_Water_1" ~
                                                                                 "ROMN_GRKO_S01_20190526_Water_1", # date adjustment
                                                                               Activity_ID == "ROMN_GRKO_S01_20190326_Sediment_0" ~
                                                                                 "ROMN_GRKO_S01_20190526_Sediment_0", # date adjustment
                                                                               Activity_ID == "ROMN_ROMO_S02_20190508_Water_0" ~
                                                                                 "ROMN_ROMO_S02_20190507_Water_1", # date adjustment
                                                                               Activity_ID == "ROMN_GLAC_022_20190930_Water_0" ~
                                                                                 "ROMN_GLAC_022_20190822_Water_1", # seems they manually adjusted
                                                                               Activity_ID == "ROMN_GLAC_022_20190930_Water_1" ~
                                                                                 "ROMN_GLAC_022_20190930_Water_0", # seems they manually adjusted
                                                                               TRUE ~ Activity_ID
                                                                               )) %>%
                                                filter(Additional_Location_Info != "GLAC_022_2019_3" & Custody_ID != "19NPS0000568 E6.FAC"), # since we merged two events we needed to remove the event that does not show up in the old data
                                              Results = edd_2019$Results  %>%
                                                mutate(Activity_ID = str_replace(Activity_ID, "_T_", "_")) %>%
                                                mutate(Activity_ID = case_when(Activity_ID == "ROMN_GRKO_S01_20190326_Water_0" ~ 
                                                                                 "ROMN_GRKO_S01_20190526_Water_0", # date adjustment
                                                                               Activity_ID == "ROMN_GRKO_S01_20190326_Water_1" ~
                                                                                 "ROMN_GRKO_S01_20190526_Water_1", # date adjustment
                                                                               Activity_ID == "ROMN_GRKO_S01_20190326_Sediment_0" ~
                                                                                 "ROMN_GRKO_S01_20190526_Sediment_0", # date adjustment
                                                                               Activity_ID == "ROMN_ROMO_S02_20190508_Water_0" ~
                                                                                 "ROMN_ROMO_S02_20190507_Water_1", # date adjustment
                                                                               Activity_ID == "ROMN_GLAC_022_20190930_Water_0" ~
                                                                                 "ROMN_GLAC_022_20190822_Water_1", # seems they manually adjusted
                                                                               Activity_ID == "ROMN_GLAC_022_20190930_Water_1" ~
                                                                                 "ROMN_GLAC_022_20190930_Water_0", # seems they manually adjusted
                                                                               TRUE ~ Activity_ID
                                                                               ))))
```

Latitude 3 
Longitude 4 
HUC8_Code 1 
HUC12_Code 1 

Activity_Type 3 
Activity_Start_Date 4 
Additional_Location_Info 1 --> they manually adjusted this?
Activity_Comment 48

Result_Detection_Condition 2 
Result_Text 2
Result_Unit 1 --> look into this
Result_Qualifier 268
Result_Status 675 
Result_Type 128 
Result_Comment 675 
Lower_Quantification_Limit 127 
Analysis_Start_Date 675 
Analysis_Start_Time 675 
Lab_Reported_Result 263
Reportable_Result 266 
